---
title: "Parameter Estimation - Global Wild Fish Catch"
author: "Genevieve Chiong"
date: "3/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(boot)
library(gt)
library(patchwork)
library(broom)
library(nlstools)
```


```{r}
# Read in the data
fish_catch <- read_csv(here("data", "fish_catch.csv")) %>% 
  rename("year" = "World Wild Fish Catch and Farmed Fish Production, 1950-2012",
         "wild_catch_mill_t" = "X2",
         "farmed_fish" = "X3", 
         "total_fish_production" = "X4") %>% 
   slice(-c(1, 2, 3, 4))

# Remove empty rows at bottom of dataset
fish_catch_na <- na.omit(fish_catch)

# Convert the year column class from character to numeric 
fish_catch_numeric <- as.data.frame(apply(fish_catch_na, 2, as.numeric))

# Create a new column with years starting at 0 instead of 1950 
fish_catch_clean <- fish_catch_numeric %>% 
  mutate(time = (year - 1950))
```


```{r}
# Exploratory graph of wild catch over time
ggplot(data = fish_catch_clean,
       aes(x = time,
           y = wild_catch_mill_t)) +
  geom_point() +
  theme_minimal() +
  labs(x = "Time (years)", 
       y = "Wild Fish Catch (million tons)")

# Log transformed data
ggplot(data = fish_catch_clean, 
       aes(x = time, 
           y = log(wild_catch_mill_t))) +
  geom_point() +
  theme_minimal() +
  labs(x = "Time (years)", 
       y = "ln(wild fish catch)")
```

```{r}
# Estimate the growth rate constant (k) *during exponential growth phase 
fish_catch_clean_exp <- fish_catch_clean %>% 
  filter(time < 20) %>% 
  mutate(ln_wild_catch = log(wild_catch_mill_t))

# Model linear to get *k* estimate (the slope of this linear equation is an estimate of the growth rate constant) 
lm_k <- lm(ln_wild_catch ~ time, data = fish_catch_clean_exp)
lm_k # Coefficient (k) ~ 0.058
```

The trend for the global wild fish catch from 1950 - 2012 follows logistic growth. The equation for logistic growth is $P(t)=\frac{K}{1+Ae^{-kt}}$, where

- $K$ is the carrying capacity
- $A$ is $\frac{K-P_0}{P_0}$
- $k$ is the growth rate constant

The initial estimates for the parameters in the model are $K$ = 93 , $A$ = 15, and $k$ = 0.058. 

```{r}
# Nonlinear least squares
fish_catch_nls <- nls(wild_catch_mill_t ~ K/(1 + A*exp(-r*time)),
                      data = fish_catch_clean,
                      start = list(K = 93, A = 15, r = 0.058),
                      trace = TRUE)

# Model summary (null hypothesis: parameter value = 0)
summary(fish_catch_nls)

# Use broom::functions to get model outputs in tidier format
model_out <- broom::tidy(fish_catch_nls)
fish_catch_nls
```
The parameters for the model describing wild catch is $$P(t) = \frac{100.28}{1+4.32e^{-0.07t}}$$

where

- $K$ = 100.28 million tons of fish
- $A$ = 4.32 
- $k$ = 0.07 million tons of fish / year


```{r}
# Make predictions for the population at all of those times (time) in the original data frame 
p_predict <- predict(fish_catch_nls)

# Bind predictions to original data frame 
fish_catch_complete <- data.frame(fish_catch_clean, p_predict)

# Plot them all together to visualize model over original observed values 
ggplot( data = fish_catch_complete,
        aes(x = time,
            y = wild_catch_mill_t)) +
  geom_point(color = "orange") + 
  geom_line(aes(x = time,
                y = p_predict),
            color = "cadetblue4") + 
  labs(x = "\nTime (years)", 
       y = "Wild Fish Catch \n(millions of tons)\n") +
  ylim(0, 105) +
  xlim(0, 63) +
  theme_minimal() 
```


### Citations 

Global wild fish catch and aquaculture production, compiled by Earth Policy Institute with 1950-2010 from U.N. Food and Agriculture Organization (FAO), Global Capture Production and Global Aquaculture Production, electronic databases, at www.fao.org/fishery/topic/16140/en.